---
title: "*C. difficile* transmission with vaccination"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(deSolve)
```

Stephenson et al. (2017) developed and analyzed an ODE model for *C. difficile* transmission with vaccination:

$$
  \begin{aligned}
    R' &= a_R\delta N - \alpha R + \theta S- k_RR\\
    S' &= a_S\delta N + \alpha R - \theta S - \lambda S + p\varepsilon D - vS - k_SS\\
    {C^-}' &= a_{C^-}\delta N - \phi C^- + (1-f)\lambda S - vC^- - k_{C^-}C^-\\
    {C^+}' &= a_{C^+}\delta N + f\lambda S - vC^+ - k_{C^+}C^+\\
    D' &= a_D\delta N + \phi C^- - p\varepsilon D - k_DD\\
    V' &= a_V\delta N + vS - \lambda V - k_VV\\
    {C_v^+}' &= a_{C_v^+}\delta N + \lambda V + vC^- + vC^+ - k_{C_v^+}C_v^+
  \end{aligned}
$$

where

$$\lambda=\beta_C(C^- + C^+ + C_v^+) + \beta_DD$$

and $a_*$ are admission class proportions and $k_*$ are discharge rates chosen to maintain a constant total population size.

For example, setting $v=0$ and choosing the baseline values (see Fig. 3 in Stephenson et al.), we have the following:

```{r, cache = FALSE}
cdiff_vacc <- function(t, state, parameters) {
  with(as.list(c(state, parameters)),
        {
          deltaN <- kR*R + kS*S + kCn*Cn + kCp*Cp + kD*D + kV*V + kCv*Cv
          lambda <- betaC*(Cn+Cp+Cv) + betaD*D
          dR <- aR*deltaN + theta*S - (alpha+kR)*R
          dS <- aS*deltaN + alpha*R + p*eps*D - (theta+lambda+v+kS)*S
          dCn <- aCn*deltaN + (1-f)*lambda*S - (phi+v+kCn)*Cn
          dCp <- aCp*deltaN + f*lambda*S - (v+kCp)*Cp
          dD <- aD*deltaN + phi*Cn - (p*eps+kD)*D
          dV <- aV*deltaN + v*S - (lambda+kV)*V
          dCv <- aCv*deltaN + lambda*V + v*(Cn+Cp) - kCv*Cv
          list(c(dR, dS, dCn, dCp, dD, dV, dCv))
        })
}

parameters <- c(
  aR = 0.75, aS = 0.09, aCn = 0.06, aCp = 0.09, aD = 0.01, aV = 0.0, aCv = 0.0,
  kR = 0.33, kS = 0.15, kCn = 0.15, kCp = 0.15, kD = 0.068, kV = 0.15, kCv = 0.15,
  alpha = 0.5, theta = 0.033, phi = 0.06, p = 0.8, eps = 0.10,
  betaC = 0.000001, betaD = 0.000001, f = 0.6, v = 0
)
state <- c(R = 112, S = 14, Cn = 9, Cp = 13, D = 2, V = 0, Cv = 0)
t <- seq(0, 30, length = 1000)

cdiff_vacc_df <- ode(y = state, time = t, func = cdiff_vacc, parms = parameters) %>%
  unclass() %>% 
  as_tibble() %>%
  mutate(
    sumC = Cn + Cp + Cv,
    sumV = V + Cv,
    deltaN = parameters["kR"]*R + parameters["kS"]*S + parameters["kCn"]*Cn + parameters["kCp"]*Cp + parameters["kD"]*D + parameters["kV"]*V + parameters["kCv"]*Cv
    ) %>%
  rename(days = time) %>% 
  gather(state, num, -days) %>% 
  mutate(state = factor(state, levels = c("R", "S", "Cn", "Cp", "sumC", "D", "V", "Cv", "sumV", "deltaN")))

cdiff_vacc_df %>% 
  filter(!state %in% c("Cp","V","Cv","deltaN")) %>% 
  ggplot(aes(days, num)) + 
    geom_line() +
    facet_wrap(~state, nrow = 3, scales = "free_y")
```

Compare two different parameter groups:

```{r, cache = FALSE}
parameters_A <- c(
  aR = 0.75, aS = 0.09, aCn = 0.06, aCp = 0.09, aD = 0.01, aV = 0.0, aCv = 0.0,
  kR = 0.33, kS = 0.15, kCn = 0.15, kCp = 0.15, kD = 0.068, kV = 0.15, kCv = 0.15,
  alpha = 0.5, theta = 0.033, phi = 0.06, p = 0.8, eps = 0.10,
  betaC = 0.000001, betaD = 0.000001, f = 0.6, v = 0
)
parameters_B <- c(
  aR = 0.75, aS = 0.09, aCn = 0.06, aCp = 0.09, aD = 0.01, aV = 0.0, aCv = 0.0,
  kR = 0.33, kS = 0.15, kCn = 0.15, kCp = 0.15, kD = 0.068, kV = 0.15, kCv = 0.15,
  alpha = 0.5, theta = 0.033, phi = 0.06, p = 0.8, eps = 0.10,
  betaC = 0.000001, betaD = 0.000001, f = 0.6, v = 0.3
)
df <- list(A = parameters_A, B = parameters_B) %>% 
      map(~ode(y = state, time = t, func = cdiff_vacc, parms = .) %>%
            unclass() %>% 
            as_tibble() %>% 
            mutate(
              sumC = Cn + Cp + Cv,
              sumV = V + Cv
            ) %>% 
            rename(days = time) %>% 
            gather(state, num, -days) %>% 
            mutate(state = factor(state, levels = c("R", "S", "Cn", "Cp", "sumC", "D", "V", "Cv", "sumV")))
            ) %>%
      bind_rows(.id = "parameter_group")
df %>% 
  filter(!state %in% c("Cp","V","Cv","deltaN")) %>% 
  ggplot(aes(days, num, col = parameter_group)) + 
    geom_line() +
    facet_wrap(~state, nrow = 3, scales = "free_y")
```

An awkward but better way to do previous:

```{r}
get_new_par_val_vec <- function(x) {
  for (i in seq_along(x)) {
    names(x)[i] <- as.character(x[i])
  }
  x
}
get_new_param_vec <- function(x, y, p) {
  y[p] <- as.numeric(x)
  y
}
parameters <- c(
  aR = 0.75, aS = 0.09, aCn = 0.06, aCp = 0.09, aD = 0.01, aV = 0.0, aCv = 0.0,
  kR = 0.33, kS = 0.15, kCn = 0.15, kCp = 0.15, kD = 0.068, kV = 0.15, kCv = 0.15,
  alpha = 0.5, theta = 0.033, phi = 0.06, p = 0.8, eps = 0.10,
  betaC = 0.000001, betaD = 0.000001, f = 0.6, v = 0
)

vary_par <- "v"
par_vals <- c(0.0, 0.3, 0.6)
par_vals <- get_new_par_val_vec(par_vals)
par_vals %>% 
  map(get_new_param_vec, parameters, vary_par) %>% 
  map(~ode(y = state, time = t, func = cdiff_vacc, parms = .) %>%
        unclass() %>% 
        as_tibble() %>% 
        mutate(
          sumC = Cn + Cp + Cv,
          sumV = V + Cv
        ) %>% 
        rename(days = time) %>% 
        gather(state, num, -days) %>% 
        mutate(state = factor(state, levels = c("R", "S", "Cn", "Cp", "sumC", "D", "V", "Cv", "sumV")))
        ) %>%
  bind_rows(.id = "par_vals") %>% 
  mutate(par_vals = as.factor(par_vals)) %>% 
  filter(!state %in% c("Cp","V","Cv","deltaN")) %>% 
  ggplot(aes(days, num, col = par_vals)) + 
    geom_line() +
    facet_wrap(~state, nrow = 3, scales = "free_y") +
    scale_color_discrete(name=vary_par)

```

Letting $v=v(t)$ be a control, we formulate an optimal control problem that seeks to minimize both cost and disease impact. 